Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
4:30 push

@@

\begin{algorithm}[t]
    \KwIn{Temporal discretization $t_n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
    \KwIn{Coarse scheme $g_{\Delta t}$}
    \KwIn{Finer scheme $g_{\textnormal{fine}}$}
    Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$\;
    Compute the corrections $\delta g(u^1_n) = g_{\textnormal{fine}}(t_n,
    u^1_n) - g_{\Delta t}(t_n, u^1_n)$ in parallel\;
    Add the prediction and correction terms as $u^2_{n+1} = g_{\Delta t}(t_n,
    u^2_n) + \delta g(u^1_n)$\;
    Repeat steps 2 and 3, incrementing the iteration label and using $u^{k+1}_0
    = u^1_0$ as the initial condition\;
 \caption{Parareal}
 \label{alg:parareal}
\end{algorithm}

@@

\subsection{Visualization of Parareal Scheme}

bksim TODO put a graph because maybe it'll help? the slides have a version

@@

\begin{equation}
\begin{aligned}
t + k\left( t + \frac{Qt}{N} \right) &< Qt \\
k &< \frac{Q - 1}{1 + Q/N} \\
N &> \frac{Qk}{Q - 1 - k} \\
\end{aligned}
\end{equation}

@@

@@

@@

@@

>>>>

\subsubsection{Speedup}

Speedup, $S$ is defined simply as the ratio of the serial implementation to the parallelized version with a given number of processors.  These numbers are usually reported per number of processors that are tested:

\begin{equation}
S = \frac{T_\textnormal{serial}}{T_\textnormal{parallel}}
\end{equation}

\subsubsection{Strong Scaling Efficiency}

Strong scaling 
Programs that are difficult to parallelize will have efficiency expressions like $\frac{1}{\ln p}$ which approaches zero as the number of processors increases by a large amount.

\begin{equation}
\textnormal{Eff}_{\textnormal{strong}} = \frac{t_1}{N * t_N} * 100\%
\end{equation}

\subsubsection{Weak Scaling Efficiency}

\begin{equation}
\textnormal{Eff}_{\textnormal{weak}} = \frac{t_1}{t_N} * 100\%
\end{equation}

@@

\begin{algorithm}
    \KwIn{Temporal discretization $t_n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
    \KwIn{Coarse scheme $g_{\Delta t}$}
    \KwIn{Finer scheme $g_{\textnormal{fine}}$}
    Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$\;
    Compute the corrections $\delta g(u^1_n) = g_{\textnormal{fine}}(t_n,
    u^1_n) - g_{\Delta t}(t_n, u^1_n)$ in parallel\;
    Add the prediction and correction terms as $u^2_{n+1} = g_{\Delta t}(t_n,
    u^2_n) + \delta g(u^1_n)$\;
    Repeat steps 2 and 3, incrementing the iteration label and using $u^{k+1}_0
    = u^1_0$ as the initial condition\;
 \caption{Parareal}
 \label{alg:parareal}
\end{algorithm}

@@

@@

\[
\begin{aligned}
t + k\left( t + \frac{Qt}{N} \right) &< Qt \\
k &< \frac{Q - 1}{1 + Q/N} \\
N &> \frac{Qk}{Q - 1 - k} \\
\end{aligned}
\]

@@

For the parallelism by space, we are dividing our grid on which we which to solve our wave diffusion equation.

@@

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$

@@

The Parareal method TODO does this scale well?

@@

Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of

\[
est
\]
